Parties, Models, Mobsters\newline {\fontsize{11}{15}\selectfont A New Implementation of Model-Based Recursive Partitioning in R}

### 2014-03-17 pkg/partykit/inst/NUS-2014/modelparty.Rnw
further improvements and touch-ups
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@491 255f4249-2263-4417-b1ea-8498779e27a7

\textbf{Example:} Logistic regression, assessing differences in
the effect of ``preferential treatment'' (``women and children first''?).

@@

\textbf{Parameter estimation:} $\widehat \theta$ by optimization of
objective function $\Psi(y, x, \theta)$ for $n$ observations $y_i$ ($i = 1, \dots, n$):

@@

\textbf{Basis:} Empirical fluctuation process of cumulative deviations
w.r.t.\ to an ordering $\sigma(z_{ij})$.

\[
W_j(t, \widehat \theta)
  \quad = \quad {\widehat B}^{-1/2} n^{-1/2}
                \sum_{i = 1}^{\lfloor nt \rfloor}
		\psi(y_{\sigma(z_{ij})}, x_{\sigma(z_{ij})}, \widehat \theta)
\qquad (0 \le t \le 1)
\]

@@

\textbf{Model:} Paired comparison via Bradley-Terry.
\begin{itemize}
  \item Paired comparisons of attractiveness for
    \emph{Germany's Next Topmodel~2007} finalists:
    Barbara, Anni, Hana, Fiona, Mandy, Anja.
  \item Survey with 192~respondents at Universit{\"a}t T{\"u}bingen.
  \item Available covariates: Gender, age, familiarty with the TV~show.
  \item Familiarity assessed by yes/no questions:
    (1)~Do you recognize the women?/Do you know the show?
    (2)~Did you watch it regularly?
    (3)~Did you watch the final show?/Do you know who won?
\end{itemize}

@@

\textbf{Here:} Use \code{mob()} directly to build model from scratch.

>>>>

\textbf{Example:} Logistic regression, assessing differences in
the effect of ``preferential treatment'' (``women and children first''?)
in the Titanic survival data.

@@

\textbf{Parameter estimation:} $\widehat \theta$ by optimization of
additive objective function $\Psi(y, x, \theta)$ for $n$ observations $y_i$ ($i = 1, \dots, n$):

@@

\textbf{Basis:} Empirical fluctuation process of cumulative deviations
w.r.t.\ to an ordering $\sigma(z_{ij})$.
\[
W_j(t, \widehat \theta)
  \quad = \quad {\widehat B}^{-1/2} n^{-1/2}
                \sum_{i = 1}^{\lfloor nt \rfloor}
		\psi(y_{\sigma(z_{ij})}, x_{\sigma(z_{ij})}, \widehat \theta)
\qquad (0 \le t \le 1)
\]

@@

\textbf{Data:} Paired comparisons of attractiveness.
\begin{itemize}
  \item \emph{Germany's Next Topmodel~2007} finalists:
    Barbara, Anni, Hana, Fiona, Mandy, Anja.
  \item Survey with 192~respondents at Universit{\"a}t T{\"u}bingen.
  \item Available covariates: Gender, age, familiarty with the TV~show.
  \item Familiarity assessed by yes/no questions:
    (1)~Do you recognize the women?/Do you know the show?
    (2)~Did you watch it regularly?
    (3)~Did you watch the final show?/Do you know who won?
\end{itemize}

\textbf{Model:} Bradley-Terry (or Bradley-Terry-Luce) model.
\begin{itemize}
  \item Standard model for paired comparisons in social sciences.
  \item Parametrizes probability $\pi_{ij}$ for preferring object $i$ over $j$
    in terms of corresponding ``ability'' or ``worth'' parameters $\theta_i$.
    \begin{eqnarray*}
          \pi_{ij}\phantom{)}  & = & \frac{\theta_i}{\theta_i + \theta_j} \\
      \mathsf{logit}(\pi_{ij}) & = & \log(\theta_i) - \log(\theta_j)
    \end{eqnarray*}
  \item Maximum likelihood as a logistic or log-linear GLM.
\end{itemize}

@@

\textbf{Here:} Use \code{mob()} directly to build model from scratch using \code{btReg.fit()} from \emph{psychotree}.
